---
title: "A Brief Cookbook for Module Writers"
author: "Zhengjia Wang"
date: '`r Sys.Date()`'
output: word_document
---

```{r, echo=F, eval=FALSE, include=FALSE}
"rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Brief Cookbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}"


"
TODO:

1. How to return values to SUMA (Push to SUMA)
2. Write a module that deals with real data
3. Async execution
4. ...
"
```

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rave)
require(shiny)
require(plotly)
require(tidyverse)
require(magrittr)
require(stringr)
require(fields)
require(uuid)

pkgname = 'RAVE'
```

## Package Structure

It's important to know the structure of `r pkgname` before writing a module. Since ECoG data is big, and we implement our package in a way that all the data is loaded into memory (though new format is under test for computers with small RAM, it's still recommeded that those computers upgrade RAM), we design the structure that seperate data repository with modules, which brings the following advantages:

* Data is shared across the sessions.
    + For example, if you open two browser tabs (running on same `r pkgname` instance), entering the IP address of the instance running. You can set different parameters for each sessions and compare the results. However, since they are running on one `r pkgname` instance, RAM shouldn't change too much.
* Data is be shared across the modules.
    + No need to import data each time a new module is opened.
    + Loading data doesn't require `shiny`, therefore we can perform analysis/writing modules without creating `r pkgname` instances. This feature will be introduced later
    + With same format for different subjects, module writers can focus on implementing visualizations. 

We are not going any further into the implementation. However, just remember that modules are session-based, and they are separated from data. When writing modules, we can always assume that the data is loaded. Therefore it's important to know what information is loaded before adding modules.

Let's load a subject without lauching any browsers. This step helps us understand setting up develop environment. However, it should never appear in the module scripts since `r pkgname` will automatically run the code for us.

```{r load_subject}
# Play with a toy dataset

# STEP 1: Set options
rave_options(
  data_dir = system.file('example/data', package = 'rave')
) -> opt;  # Assign to avoid printing


# STEP 2: Load subject
find_subject_ids()   # List all subject IDs
subject_id = 'dipterix' # Let's use the first ID as an example

subject = get_subject(subject_id)  # Load subject meta-data


# STEP 3: Load ECoG value, for example, first four
subject$electrode_label_by_index(1:4)
suppressMessages({   # Suppress loading messages.
  subject$data_environment$load(electrodes = 1:4)             # Prepare electrodes
  subject$data_environment$bind_electrodes(electrodes = 1:4, debug = T)  # Concatenate tensors
})   


# Check data
data_env = subject$data_environment
dim(data_env$data) # Trial x Freq x Time x Electrode (Valid)

rm(subject, data_env)
```

During module development, we use function `attach_virtualenv` to load subject data. In this case, we can also use `attach_virtualenv(subject_id, 1:4)` after step 1 to get `data_env`. However, `attach_virtualenv` hides some details.

We can devide the code above into three steps as follows:

1. Setting up option, or creating configuration file
2. Load a subject's meta data into memory
3. Load electrodes needed for the analysis

(*Read vignette `A Cookbook for `r pkgname` Users` for details in changing settings and preparing subject data.*)

Here we can see that in order to load data environment, `r pkgname` applies two steps. First, loading subjects' meta information, which is in `[subject]/ecog/meta`, the descriptions of trials, frequencies, time points and electrodes. Then, according to user's requirement, load electrodes of interest.

The result in this case, `data_env`, has one element `data`, which is a four-mode tensor: `trial`x`frequency`x`time`x`electrode`. If we apply `attach_virtualenv`, we can see other elements such as `cumsum`, `electrodes`, etc. 

```{r attach_virtualenv}
attach_virtualenv(
  subject_id = subject_id,
  electrodes = 1:4
)
ls(data_env)
```

- `data_env$data`: four-mode tensor
- `data_env$cumsum`: four-mode tensor, cumulative summation of `data_env$data` over time
- `data_env$electrodes`: current loaded electrodes
- `data_env$subject`: subject object that contains meta data
- `get_global_var`: see next part

Now we have created environment for writing modules.

## Writing the first module


### Structure of a basic module

To write our first module, here's what we have:

> `data_env$data`, `data_env$cumsum`, `data_env$electrodes`, `data_env$subject`

These data are **read-only**. Never change them. Also, we don't recommend assignments such as `mydata <- data_env$data` since it will copy the data, which could be as large as 10GB, in the memory.

To make it a module, here's what we need to implement:

> `SHINY_INPUT`, `SHINY_EXECUTE`, `SHINY_OUTPUT`

There are some other variables that are optional:

> `SHINY_DESCRIPTION`

Please copy this template save save it as a `example.R` file.

```
###########################
# Example module
# Author: Zhengjia Wang
###########################


# Definitions of input
# There are more than seven types of inputs, for "text_input", type "?text_input" to see documentations
#
# For all inputs, "inputId" are important since they are local names for your inputs
# For example, "text_input(inputId = 'textA',..."
# Here we have a variable "params$textA" that you can use as a *string* in `SHINY_EXECUTE`
SHINY_INPUT = list(
  text_input(inputId = 'textA', label = 'Please enter a text', init = function(){
    return(list(
      value = get_local_var('textA', paste('Default text', Sys.time()))
    ))
  }),
  numeric_input(inputId = 'numB', label = 'This is a random number', value = 0, init = function(){
    return(list(
      value = rnorm(1)
    ))
  })
)

# (optional) parameters for local debug
# We have two input IDs, "textA", and "numB"
params = list(
  textA = 'debug text',
  numB = 1
)

# Algorithm part, process data here
SHINY_EXECUTE = function(params, ...){
  
  # Pre-process data
  
  s = paste(params$textA, '|', params$numB)
  
  # Return a named list of functions
  return(list(
    
    # The name "output_text" will be used as output ID in "SHINY_OUTPUT"
    output_text = function(){
      return(s)
    }
  ))
}

# Visualization settings, a named list of output tabs and components
SHINY_OUTPUT = list(
  
  # "First tab" is the name of the tab
  `First tab` = list(
    
    # "output_text" is output ID defined in "SHINY_EXECUTE"
    verbatimtext_output(outputId = 'output_text', title = 'Output Text')
  ),
  `Second tab` = list()
)
```

This is our first module. It will verbatim a text which concatenate the two inputs. However, in order to put them in `r pkgname`, we need two more steps: 

1. Create a module index csv file (for example, `modules.csv`), and enter/append the following table. Check `system.file('modules.csv', package = 'rave')` for example

```{r, echo=FALSE}
# This is an example file from 
module_index = read.csv(
  system.file('example/module/modules.csv', package = 'rave'),
  stringsAsFactors = F
)
module_index$source_path = 
  eval(parse(text = module_index$source_path))

knitr::kable(module_index)
```

2. Register index file `modules.csv` to `r pkgname`:

```
rave_options(
  module_lookup_file =
    '`r system.file('example/module/modules.csv', package = 'rave')`'
)
```

Now let's lauch the web UI
```
# Set options,
# Specify data directory
# Specify module index file
# 
rave_options(
  data_dir = system.file('example/data', package = 'rave'),
  module_lookup_file = system.file('example/module/modules.csv', package = 'rave')
) -> opt;

# Command to lauch app
init_app()
```


### Inputs

`SHINY_INPUT` is a list of input components. Here is the list for inputs:

```{r, echo=FALSE}

data_frame(
  Component = c('text_input', 'numeric_input', 'checkbox_input', 'select_input', 'slider_input', 'action_button', 'file_input'),
  Type = c('Characters', 'Number', 'Boolean', 'Characters', 'Number(s)', 'Signal', 'File Table')
) %>% 
  knitr::kable()


```

Usage:


```
text_input(inputId, label, init, global_var, value, width, placeholder)
numeric_input(inputId, label, init, global_var, value, min, max, step, width)
checkbox_input(inputId, label, init, global_var, value, width)
select_input(inputId, label, init, global_var, choices, selected, multiple, width, size)
slider_input(inputId, label, init, global_var, min, max, value, 
             step, round, format, locale, ticks, width, sep, pre,
             post, timeFormat,timezone, dragRange)
action_button(inputId, label, init, global_var, icon, width)
file_input(inputId, label, init, global_var, multiple, accept,
           width, buttonLabel, placeholder)
```

* `inputId`: (Mandatory) variable name, will be stored in `params` and called as `params$inputId` or `params[['inputId']]`
* `label`: (Mandatory) text that will be displayed in app
* `value/selected`: (some are optional), default value for components
* `init`: (optional) a function with no arguments and returns named list of initial/default values for components
* `global_var`: (optional) name for the component if this input will be used in other modules. `NULL` by default, meaning that the component can only be accessed by local module.
* `width`: (optional) width of the component, usually ignore this argument
*(Type "?" for documentation, for example, `?text_input`)*

### Data process

`SHINY_EXECUTE` is the body part that we process our data with inputs. The format follows:

```
1  SHINY_EXECUTE = function(params, ...){
2    [YOUR_CODE_HERE]
3   
4    return(list(
5      [OUTPUT_ID_1] = function(){
6        [CODE_GENERATING_GRAPH/TABLE]
7      },
8      [OUTPUT_ID_2] = function(){
9       ...
10     },
11     ...
12   ))
13 }
```

Line `1` is fixed. We can see that `SHINY_EXECUTE` itself is a R function, which takes `params` and `...` as arguments (`...` is a special keyword, reserved for future development).

`params` is a list, with number of elements equaling to the number of inputs, and their names are input IDs. For instance, if `SHINY_INPUT` has a component with `inputId='textinput'`, then we can access users' inputs for this component via `params$textinput` or `params[['textinput']]` within `SHINY_EXECUTE`.

The output of `SHINY_EXECUTE` should be a list of functions, with output ID as the name for each functions. For example, if we want to have a plot output with `outputId='myplot'` which plots `Sin(x)` in the `SHINY_OUTPUT`, then we need to have `SHINY_EXECUTE` returns a list as the example below:

```
[WITHIN_SHINY_EXECUTE]
...
return(list(
  myplot = function(){
    plot.function(sin)
  }
))
[END_OF_SHINY_EXECUTE]
```

Make sure that `myplot` is a function that takes no arguments.


### Outputs

`SHINY_OUTPUT` defines output. The structure of `SHINY_OUTPUT` is defined as a list of tabs, with each tab containing output components.

```
tabcomponets_1 = list(
  [Output Component 1],
  [Output Component 2],
  ...
)

...

SHINY_OUTPUT = list(
  [Tab name 1] = tabcomponets_1,
  [Tab name 2] = tabcomponets_2,
  ...
)
```

The output will be displayed in tabs, with "[Tab name 1]", "[Tab name 2]", etc. Within each tabs ("tabcomponets_1", "tabcomponets_2", ...), there should be a list of components. Similar to the inputs, there are four types of outputs:

> verbatimtext_output, datatable_output, plot_output, plotly_output

Usage:

```
***_output(outputId, title, ...)
```

* `outputId`: matches with names of function returned by `SHINY_EXECUTE`
* `title`: title for output to be displayed

Here's an example for plots

```
SHINY_EXECUTE = function(params, ...){
  ...
  
  return(list(
    plotA = function(){
      plot.function(sin)
    }
  ))
}

SHINY_OUTPUT = list(
  Visualization = list(
    plot_output(outputId = 'plotA', title = 'Curve for Sin(x)')
  )
)
```


